Algebraic Groups\vspace{-1ex}

### 2019-10-16 Algebraic Groups/Algebraic-Groups.tex
update

	The second follows by considering $v\in V$ and looking at the image of it under $\sigma$. That is, 
	\[\sigma(v)=\sum_1^Nt^{\alpha_i}\otimes v_i\]
	where $\alpha_i\in A$ and $v_i\in V$. Then a very simple argument shows $v=\sum v_i$ (using linearity).
	Then it remains to show that $v_i\in V_{a_i}$, but this will make things work. (use the other axiom of a coaction).
\end{prf}
\begin{rmk}
	When $A=\bbZ$, $G=D(\bbZ)=\Gm$, which tells us that representations of $\Gm$ are in bijection with $\bbZ$-gradings 
	of $V\cong\oplus_{n\in\bbZ}V_n$!
\end{rmk}
\begin{defn}
	A linear algebraic group $G\to\Spec k$ is called \textbf{linear reductive} if every representation decomposes as a direc sum of irreducibles.
\end{defn}
\begin{prob}
	Show the above is equivalent to the statements
	\begin{itemize}
		\item for each $G$-representations $W\subseteq V$, there exists $W'\subseteq V$ subrepresentations such that $V\cong W\oplus W'$.
		\item $0\to W\to V\to W'\to 0$ is exact.
	\end{itemize} 
\end{prob}
\begin{rmk}
	Notice that this says that $D(A)$ is linear reducible. In particular, $\Gm$ and $|mu_n$ are in \textbf{any characteristic}. This runs counter to Maschke in 
	finite groups.
\end{rmk}

@@

As another example, consider
\[\Ga\cong\left\{\begin{pmatrix}
	1 & \alpha \\ 0 & 1
\end{pmatrix}\right\}
\subset\GL_2(k)\]
where $\Ga$ acts on $k^2$ by $\alpha(x,y)=(a+\alpha y,y)$. Then it can be easily seen not to be alinear representation.

>>>>

	The second follows by considering $v\in V$ and looking at the image of it under $\sigma$. That is, 
	\[\sigma(v)=\sum_1^Nt^{\alpha_i}\otimes v_i\]
	where $\alpha_i\in A$ and $v_i\in V$. Then a very simple argument shows $v=\sum v_i$ (using linearity).
	Then it remains to show that $v_i\in V_{a_i}$, but this will make things work. (use the other axiom of a coaction).
\end{prf}
\begin{rmk}
	When $A=\bbZ$, $G=D(\bbZ)=\Gm$, which tells us that representations of $\Gm$ are in bijection with $\bbZ$-gradings 
	of $V\cong\oplus_{n\in\bbZ}V_n$!
\end{rmk}
\begin{defn}
	A linear algebraic group $G\to\Spec k$ is called \textbf{linear reductive} if every representation decomposes as a direct sum of irreducibles.
\end{defn}
\begin{prob}
	Show the above is equivalent to the statements
	\begin{itemize}
		\item for each $G$-representations $W\subseteq V$, there exists $W'\subseteq V$ subrepresentations such that $V\cong W\oplus W'$.
		\item $0\to W\to V\to W'\to 0$ is exact.
	\end{itemize} 
\end{prob}
\begin{rmk}
	Notice that this says that $D(A)$ is linear reducible. In particular, $\Gm$ and $|mu_n$ are in \textbf{any characteristic}. This runs counter to Maschke in 
	finite groups.
\end{rmk}

@@

As another example, consider
\[\Ga\cong\left\{\begin{pmatrix}
	1 & \alpha \\ 0 & 1
\end{pmatrix}\right\}
\subset\GL_2(k)\]
where $\Ga$ acts on $k^2$ by $\alpha(x,y)=(a+\alpha y,y)$. Then it can be easily seen not to be a linear representation.

\section{October 16th, 2019}
Last time we talked about representations! Woot.

Notice that if $G$ is linear (i.e. affine), then the multiplication map induces comultiplication 
\[\Gamma(G)\to \Gamma(G)\otimes\Gamma(G)\]
so $\Gamma(G)$ is the \textbf{regular} representation with coaction given by multiplication.

We also saw some equivalent contitions similar to Machke for linear reductive groups. Finally we say some examples and diagonalizable groups.

\subsection{New Stuff}
Given a $G$-representation $V$, let $V^G$ be 
\[\{v\in V|\sigma(v)=1\otimes v\}=\operatorname{Eq}\{V\stackrel{\xrightarrow{1\otimes -}}{\xrightarrow{\sigma}}\Gamma(G)\otimes V\}=\Hom^G(k,V)\subseteq V.\]
\begin{rmk}
	I need to figure out the \TeX for equalizers/parallel maps.
\end{rmk}
\begin{ex}
	Given the representation $\Gm$ action on $V=\oplus V_d$, $V^G=V_0$.
\end{ex}
\begin{prob}
	If $G(k)$ is dense in $G$, then $V^G=V^{G(k)}$.
\end{prob}
\begin{prop}
	A linear algebraic group $G$ over $k$ is linearly reductive if and only if the functor from $G$-representations to 
	$k$-vector spaces given by $V\mapsto V^G$ is exact.
\end{prop}
\begin{prf}
	If $V\cong W\oplus W'\twoheadrightarrow W$ is a $G$ representation, then 
	\[W^G\oplus(W')^G=V^G\twoheadrightarrow W^G\]
	is also surjective.

	Suppose that we have a short exact sequence 
	\[0\to W'\to V\to W\to 0\]
	and that this functor is exact. Then we want to show we get a section $\sigma:W\to V$. To do this, consider the functor 
	$\Hom^G(W,-)=\Hom^G(k,W^\vee\otimes -)=(W^\vee\otimes -)^G$, so by the assumption this is exact and we can lift the identity on $W$ 
	to a map in $\Hom^G(W,V)$, giving us our section.
\end{prf}
\begin{prop}
	Let $G$ be a linear algebraic group over $k$ and $V$ a $G$-representation. Let $W\subseteq V$ be a finite dimensional $k$-subspace (not 
	necessarily $G$-invariant). Then there exists $W\subseteq W'\subseteq V$ such that $W'$ is a finite dimensional representation of $G$.
\end{prop}
\begin{prf}
	We can assume that $W=\langle w\rangle$ for $w\in V$. Apply $\sigma:V\to \Gamma(G)\otimes V$. Then if $\{t_i\}$ is a 
	basis for $\Gamma(G)$, we get 
	\[w\mapsto\sum t_i\otimes w_i.\]
	Then we claim that $w\in\langle w_i\rangle$ and $\langle w_i\rangle\subseteq V$ is a subrepresentation.

	For the first, consider the diagrams:
	\begin{center}
		\begin{tikzcd}
			k\otimes V & \Gamma(G)\otimes V\ar[l,swap,"e^\ast\otimes\id"]\\
			& V\ar[ul,"\sim"]\ar[u,"\sigma",swap]
		\end{tikzcd}
		\begin{tikzcd}
			\sum e^\ast(t_i) w_i & \sum t_i\otimes w_i\ar[l]\\
			& w\ar[u]
		\end{tikzcd}
	\end{center}
	so $w$ is in the span of the $w_i$.

	For the second claim, we need to show that 
	\[\sigma(w_i)\in\Gamma(G)\otimes\langle w_i\rangle.\]
	To see this consider the diagram 
	\begin{center}
		\begin{tikzcd}
			\Gamma(G)\otimes\Gamma(G)\otimes V & \Gamma(G)\otimes V\ar[l,"m^\ast\otimes\id"]\\
			\Gamma(G)\otimes V\ar[u,"\id\otimes\sigma"] & V\ar[l,"\sigma"]\ar[u,"\sigma"]
		\end{tikzcd}
	\end{center}
	And tracing through $w\in V$, we get that 
	\[\sum t_i\otimes\sigma(w_i)=\sum_{i,j,k}\alpha_{i,j,k}t_i\otimes t_j'\otimes w_k\]
	and so by looking at coefficients of $t_i\otimes \Gamma(G)\otimes V=\sigma(w_i)$ (look closer here), we see it is 
	\[\sum_{j,k}\alpha_{i,j,k}t_j'\otimes w-k\in\Gamma(G)\otimes\langle w_i\rangle.\]
\end{prf}
\begin{cor}
	If $V$ is a $G$ representation, then 
	\[W=\bigcup_{W\subset V}W\]
	where the union is over all finite dimensional subgroups.
\end{cor}
\begin{cor}
		If $G$ is a linear algebraic group (affine finite type over $k$), then for some $n$ $G\subseteq \GL_n$ is a closed subgroup.
		In other workds, there exists a faithful representation $V$ of $G$.
\end{cor}

Now consider the regular representation $\Gamma(G)\to \Gamma(G)\otimes \Gamma(G)$. Notice that $\Gamma(G)$ is 
a $k$-algebra of finite type. Choose generators $g_1,\dots,g_n$ for $\Gamma(G)$. Take a subrepresentation of $\Gamma(G)$
spanned by $\langle h_1,\dots,h_N\rangle$ containing the span of the $g_i$.

We have a map $G\to \GL(W)$ where $W=\langle h_i\rangle$ and we aske whether the induced map 
\[\Sym^\ast(W\otimes W^\vee)_{\det}\]
is surjective.

Let's say that $h_i\mapsto \sum_j\gamma_{i,j}\otimes h_j$ under $\sigma$. Then using this map and the natural pairing between $W$ and $W^\vee$, we get a map 
\[\Sym^\ast(W\otimes W^\vee)\supset W\otimes W^\vee\to\Gamma(G)\otimes W\otimes W^\vee\to \Gamma(G)\]
where we send 
\[h_i\otimes h_j^\ast\mapsto \gamma_{i,j}\]
So using the counit identiy we can write 
\[h_i=\sum_j e^\ast(\gamma_{i,j})h_j\]
but we really want to write $h_i$ as a linear combination of the $\gamma_{i,j}$ (since we have shown they all lie in the image of this map).
We don't know how to finish up.
